Covariate analysis of ROSMAP reprocessed counts (with TMM normalisation)

### 2017-03-07 code/Rmd/MSSM_RNASeq_Reprocessed_TMM_BinnedCDR.Rmd
Added picard metrics to MSSM reprocessed rnaseq analysis

library(parallel)
library(foreach)
library(doMC)
library(foreach)
library(doParallel)

@@

# Get technical metadata
METADATA_TECH_ID <- 'syn6100548'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_TECH_ID
METADATA_TECH <- read.table(synGet(METADATA_TECH_ID)@filePath,sep=',',header=T, row.names=1) %>%
  group_by(sampleIdentifier) %>%
  top_n(1, -rRNA.rate) %>%
  dplyr::select(sampleIdentifier, BrodmannArea, barcode, individualIdentifier, batch, RIN, TotalReads, Mapped, rRNA.rate) %>%
  unique

@@

METADATA <- full_join(METADATA_TECH, METADATA_CLINICAL)
```

@@

# Total Reads
p[[4]] = ggplot(METADATA, aes(x = BrodmannArea.Dx, y = as.numeric(TotalReads), fill = BrodmannArea)) + geom_boxplot()
p[[4]] = p[[4]] + ggtitle('Total Reads') + theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1))

# Mapped Reads
p[[5]] = ggplot(METADATA, aes(x = BrodmannArea.Dx, y = as.numeric(Mapped), fill = BrodmannArea)) + geom_boxplot()
p[[5]] = p[[5]] + ggtitle('Mapped') + theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1))

# Mapped Reads
p[[6]] = ggplot(METADATA, aes(x = BrodmannArea.Dx, y = as.numeric(rRNA.rate), fill = BrodmannArea)) + geom_boxplot()
p[[6]] = p[[6]] + ggtitle('rRNA Rate') + theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1))

@@

### Covariate clustering
Determine relationship between covariates. 
```{r covariates.clustering}
FactorCovariates <- c("individualIdentifier", "batch", "RACE", "SEX", "BrodmannArea.Dx")
ContCovariates <- c("RIN", "RIN2", "AOD","PMI", "TotalReads", "Mapped", "rRNA.rate")

@@

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.character)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.numeric)
```
Covariate correlation
```{r covariates.correlation, fig.width=8, fig.height=8}
COVARIATES.CORRELATION = CovariateAnalysis::getAssociationStatistics(COVARIATES, PVAL = 0.05)
COVARIATES.CORRELATION$plot
```
### Gene Filtering
Preprocess counts matrix and metadata. Determine design matrix for normalisation and differential expression analysis.

@@

# Find significant covariates
adjustCovars = preAdjustedSigCovars$significantCovars
```
Significant covariates to adjust at FDR 0.1 are `r paste(gsub('_','\\\\_',adjustCovars), collapse= ',')`
```{r preAdjustedSigCovars.NULL.ALL, fig.width=25, fig.height=12}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```
### Normalisation (iterative)
Since many covariates are correlated, re-normalising COVARIATES with an iterative design matrix.
```{r iterative.adjusted.voom.normalisation, results='asis'}
postAdjustCovars = c('BrodmannArea.Dx', 'batch', 'SEX')
randomCovariates = 'individualIdentifier'

# Assign residual covariates
residualSigCovars = preAdjustedSigCovars
significantCovars = residualSigCovars$significantCovars
covariatesEffects = residualSigCovars$Effects.significantCovars
covariatesEffects = covariatesEffects[setdiff(significantCovars, c(postAdjustCovars, randomCovariates))]

@@

  # Residual covariates to choose from
  residCovars <- setdiff(c(FactorCovariates,ContCovariates), c(postAdjustCovars, randomCovariates))

@@

  # Add postadjusted covariates (if any)
  significantCovars = residualSigCovars$significantCovars
  covariatesEffects = residualSigCovars$Effects.significantCovars
  covariatesEffects = covariatesEffects[setdiff(significantCovars,c(postAdjustCovars, randomCovariates))]

@@

tmp <- paste('Using following covariates in the final model:', modelStr, 'and Donor_ID as random effects')
```
`r tmp`
### Sanity Check
Residualise significant covariates using a linear model and find correlation between PCA of residuals with covariates
```{r residualSigCovars.ALL, fig.width=25, fig.height=12}
residualSigCovars[["PC_res"]][[2]]$plotData
```
Adding "BrodmannArea.Dx" back to the residuals
```{r addVarsBack}
varsToAddBackIn <-  grep("BrodmannArea.Dx", colnames(FIT$coefficients), value = T)
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  t(data.matrix(DM1$design[,varsToAddBackIn]) %*% t(FIT$coefficients[,varsToAddBackIn]))
```

@@

```{r get.gene.symbols}
backgroundGenes = data.frame(gene.id = rownames(COUNT)) %>%
  dplyr::mutate(ensembl_gene_id = gene.id) %>%
  tidyr::separate(ensembl_gene_id, c('ensembl_gene_id', 'position'), sep = '\\.')

# Define biomart object
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", dataset = "hsapiens_gene_ensembl", host = "dec2016.archive.ensembl.org")

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", 
                       values = backgroundGenes$ensembl_gene_id,
                       mart = mart)

# Add gene names
RESIDUAL.GENE_EXPRESSION <- RESIDUAL.GENE_EXPRESSION %>%
  rownameToFirstColumn('gene.id') %>%
  left_join(backgroundGenes) %>%
  left_join(Ensemble2HGNC) %>%
  dplyr::select(gene.id, ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))

VOOM.ADJUSTED.GENE_EXPRESSION$E <- VOOM.ADJUSTED.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('gene.id') %>%
  left_join(backgroundGenes) %>%
  left_join(Ensemble2HGNC) %>%
  dplyr::select(gene.id, ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp, fig.height=10, fig.width=20}
# Fit contrast
cntr = c()
for (BA in unique(METADATA$BrodmannArea)){
  tmp = combn(paste0('BrodmannArea.Dx', BA, '.', c('ND','MD','SD')),2)
  cntr = c(cntr, apply(tmp, 2, paste, collapse = '-'))
}

@@

# Get differnetial expression
DE = lapply(1:12, function(i, FIT.CONTR, VOOM.ADJUSTED.GENE_EXPRESSION){
  topTable(FIT.CONTR, coef=i, number = dim(VOOM.ADJUSTED.GENE_EXPRESSION$E)[1], confint = T)
}, FIT.CONTR, VOOM.ADJUSTED.GENE_EXPRESSION)
names(DE) = gsub('BrodmannArea.Dx','',colnames(contrast))

DE1 = plyr::ldply(DE, .id = 'Comparison') %>%
  dplyr::mutate(Comparison1 = Comparison) %>%
  tidyr::separate(Comparison1, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('BrodmannArea','From.Dx'), sep = 5) %>%
  tidyr::separate(to.state, c('BrodmannArea','To.Dx'), sep = 5) %>%
  tidyr::unite(Dx, From.Dx, To.Dx, sep = '-') %>%
  dplyr::select(-Comparison) %>%
  dplyr::rename(Region = BrodmannArea, Comparison = Dx, gene.id = genes) %>%
  dplyr::mutate(Region = factor(Region, labels = c('FP', 'STG', 'PHG', 'IFG')),
                Study = 'MSBB') %>%
  left_join(backgroundGenes) %>%
  left_join(Ensemble2HGNC) %>%
  dplyr::select(-position)

p <- ggplot(DE1, aes(x = logFC, y = -log10(adj.P.Val))) + geom_point()
p <- p + geom_hline(yintercept = -log10(0.05), color = 'red')
p <- p + geom_vline(xintercept = log2(1.2), color = 'red')
p <- p + geom_vline(xintercept = -log2(1.2), color = 'red')
p <- p + facet_grid(Region~.+Comparison)
p

@@

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = 'MSBB Reprocessed RNASeq TMM Binned CDR', parentId = parentId)
annotations(CODE) = list(
  dataType = 'mRNA',
  dataSubType = 'geneExp',
  summaryLevel = 'gene',
  assay	 = 'Reprocessed RNAseq',

@@

# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'sampleIdentifier')
write.table(COVARIATES, file = 'MSBBCovariates.tsv', sep = '\t', row.names=F, quote=F)
COV_OBJ = File('MSBBCovariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix %>%
  as.matrix() %>% rownameToFirstColumn('gene.id') %>%
  write.table(file = 'MSBBCounts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('MSBBCounts.tsv', name = 'Raw counts (filtered)', parentId = CODE$properties$id)
COUNT_OBJ = synStore(COUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                     executed = thisFile, activityDescription = activityDescription)

# Store logCPM
write.table(VOOM.ADJUSTED.GENE_EXPRESSION$E, file = 'MSBBlogCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('MSBBlogCPM.tsv', name = 'Raw counts (logCPM)', parentId = CODE$properties$id)
LCOUNT_OBJ = synStore(LCOUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)

# Store design matrix
write.table(DM1$design, file = 'MSBBDesign.tsv', sep = '\t', row.names=F, quote=F)
DM_OBJ = File('MSBBDesign.tsv', name = 'Design Matrix', parentId = CODE$properties$id)
DM_OBJ = synStore(DM_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                  executed = thisFile, activityDescription = activityDescription)

# Store residual gene expression
write.table(RESIDUAL.GENE_EXPRESSION, file = 'MSBBExpression.tsv', sep = '\t', row.names=F, quote=F)
EXP_OBJ = File('MSBBExpression.tsv', name = 'Normalised Adjusted Residual Expression (Dx added)', 
               parentId = CODE$properties$id)
EXP_OBJ = synStore(EXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)

# Store differential expression results
write.table(DE1, file = 'MSBBDiffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('MSBBDiffExpression.tsv', name = 'Differential Expression Analysis', parentId = CODE$properties$id)
DEXP_OBJ = synStore(DEXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                    executed = thisFile, activityDescription = activityDescription)

stopCluster(cl)
```
|  *Results*                |  *SynapseID*                    |
|  -------                  |   ---------                     |
|  Covariates               |  `r COV_OBJ$properties$id`      |
|  Raw counts (logCPM)      |  `r LCOUNT_OBJ$properties$id`   |
|  Design Matrix            |  `r DM_OBJ$properties$id`       |
|  Residual Expression      |  `r EXP_OBJ$properties$id`      |
|  Differential Expression  |  `r DEXP_OBJ$properties$id`     |

>>>>

library(parallel)
library(foreach)
library(doParallel)

@@

# Get technical metadata
METADATA_TECH_ID <- 'syn6100548'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_TECH_ID
METADATA_TECH <- read.table(synGet(METADATA_TECH_ID)@filePath,sep=',',header=T, row.names=1) %>%
  group_by(sampleIdentifier) %>%
  top_n(1, -rRNA.rate) %>%
  dplyr::select(sampleIdentifier, BrodmannArea, barcode, individualIdentifier, batch, RIN) %>%
  unique

@@

# Get picard metrics from synapse
METADATA.PICARD_ID <- 'syn8380379'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.PICARD_ID
METADATA.PICARD_FPATHS <- untar(synGet(METADATA.PICARD_ID)@filePath, list = FALSE, exdir = getwd())
METADATA.PICARD_FPATHS <- untar(synGet(METADATA.PICARD_ID)@filePath, list = TRUE, exdir = getwd())
METADATA.PICARD_FPATHS = METADATA.PICARD_FPATHS[grep('picard.CombinedMetrics.csv', METADATA.PICARD_FPATHS)]
METADATA.PICARD <- plyr::ldply(METADATA.PICARD_FPATHS, read.csv)

METADATA.PICARD$sampleIdentifier = gsub('picard/', '', METADATA.PICARD_FPATHS) %>%
  gsub('.accepted_hits.sort.coord/picard.CombinedMetrics.csv','',.)
colnames(METADATA.PICARD) = gsub('AlignmentSummaryMetrics__','',colnames(METADATA.PICARD))
colnames(METADATA.PICARD) = gsub('RnaSeqMetrics__','',colnames(METADATA.PICARD))

METADATA <- full_join(METADATA_TECH, METADATA_CLINICAL) %>%
  inner_join(METADATA.PICARD)
```

@@

# RIBOSOMAL FRACTION
p[[4]] = ggplot(METADATA, aes(x = BrodmannArea.Dx, y = as.numeric(PCT_RIBOSOMAL_BASES), fill = BrodmannArea)) + geom_boxplot()
p[[4]] = p[[4]] + ggtitle('Fraction of Ribosomal Bases') + theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1))

# INTRONIC BASES
p[[5]] = ggplot(METADATA, aes(x = BrodmannArea.Dx, y = as.numeric(PCT_INTRONIC_BASES), fill = BrodmannArea)) + geom_boxplot()
p[[5]] = p[[5]] + ggtitle('Frction Intronic Bases') + theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1))

# Mapped Reads
p[[6]] = ggplot(METADATA, aes(x = BrodmannArea.Dx, y = as.numeric(METADATA$PCT_INTERGENIC_BASES), fill = BrodmannArea)) + geom_boxplot()
p[[6]] = p[[6]] + ggtitle('Fraction Intergenic Bases') + theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1))

@@

### Covariate clustering
Determine relationship between covariates. 
```{r covariates.clustering}
FactorCovariates <- c("individualIdentifier", "batch", "RACE", "SEX", "BrodmannArea.Dx")
ContCovariates <- c("RIN", "RIN2", "AOD","PMI", "PCT_PF_READS", "PCT_PF_READS_ALIGNED", 
                    "PCT_INTERGENIC_BASES", "PCT_INTRONIC_BASES", 
                    "PCT_RIBOSOMAL_BASES", "PCT_USABLE_BASES")

@@

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.character)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.numeric)
```
Covariate correlation
```{r covariates.correlation, fig.width=8, fig.height=8}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES[,-(11)])
draw(COVARIATES.CORRELATION$plot, heatmap_legend_side = 'left')
```
### Gene Filtering
Preprocess counts matrix and metadata. Determine design matrix for normalisation and differential expression analysis.

@@

# Find significant covariates
adjustCovars = preAdjustedSigCovars$significantCovars
```
Significant covariates to adjust at FDR 0.1 are `r paste(gsub('_','\\\\_',adjustCovars), collapse= ',')`
```{r preAdjustedSigCovars.NULL.ALL, fig.width=25, fig.height=12}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```
### Normalisation (iterative)
Since many covariates are correlated, re-normalising COVARIATES with an iterative design matrix
```{r iterative.adjusted.voom.normalisation, results='asis'}
postAdjustCovars = c('batch', 'SEX')
randomCovariates = 'individualIdentifier'

# Assign residual covariates
residualSigCovars = preAdjustedSigCovars
significantCovars = residualSigCovars$significantCovars
covariatesEffects = residualSigCovars$Effects.significantCovars
covariatesEffects = covariatesEffects[setdiff(significantCovars, c(postAdjustCovars, randomCovariates,'BrodmannArea.Dx'))]

@@

  # Residual covariates to choose from
  residCovars <- setdiff(c(FactorCovariates,ContCovariates), c(postAdjustCovars, randomCovariates, 'BrodmannArea.Dx'))

@@

  # Add postadjusted covariates (if any)
  significantCovars = residualSigCovars$significantCovars
  covariatesEffects = residualSigCovars$Effects.significantCovars
  covariatesEffects = covariatesEffects[setdiff(significantCovars,c(postAdjustCovars, randomCovariates, 'BrodmannArea.Dx'))]

@@

tmp <- paste('Using following covariates in the final model:', modelStr, 'and Donor_ID as random effects')
```
`r tmp`
### Sanity Check
Residualise significant covariates using a linear model and find correlation between PCA of residuals with covariates
```{r residualSigCovars.ALL, fig.width=25, fig.height=12}
# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, 
                                              COVARIATES,
                                              'adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)

residualSigCovars[["PC_res"]][[2]]$plotData
```

@@

```{r get.gene.symbols, cache=TRUE, eval=T}
backgroundGenes = data.frame(gene.id = rownames(COUNT)) %>%
  dplyr::mutate(ensembl_gene_id = gene.id) %>%
  tidyr::separate(ensembl_gene_id, c('ensembl_gene_id', 'position'), sep = '\\.')

# Define biomart object
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", host = "dec2016.archive.ensembl.org", dataset = "hsapiens_gene_ensembl")

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", 
                       values = backgroundGenes$ensembl_gene_id,
                       mart = mart)
```
### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp, fig.height=10, fig.width=20}
# Post adjusted design matrix
DESIGN = getDesignMatrix(COVARIATES[,'BrodmannArea.Dx',drop=F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]

# Fit linear model with residual expression
FIT = lmFit(RESIDUAL.GENE_EXPRESSION, design = DESIGN$design)

# Fit contrast
cntr = c()
for (BA in unique(METADATA$BrodmannArea)){
  tmp = combn(paste0('BrodmannArea.Dx', BA, '.', c('ND','MD','SD')),2)
  cntr = c(cntr, apply(tmp, 2, paste, collapse = '-'))
}

@@

# Get differnetial expression
DE = lapply(1:12, function(i, FIT.CONTR, RESIDUAL.GENE_EXPRESSION){
  topTable(FIT.CONTR, coef=i, number = dim(RESIDUAL.GENE_EXPRESSION)[1], confint = T) %>%
    rownameToFirstColumn('genes')
}, FIT.CONTR, RESIDUAL.GENE_EXPRESSION)
names(DE) = gsub('BrodmannArea.Dx','',colnames(contrast))

DE1 = plyr::ldply(DE, .id = 'Comparison') %>%
  dplyr::mutate(Comparison1 = Comparison) %>%
  tidyr::separate(Comparison1, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('BrodmannArea','From.Dx'), sep = 5) %>%
  tidyr::separate(to.state, c('BrodmannArea','To.Dx'), sep = 5) %>%
  tidyr::unite(Dx, From.Dx, To.Dx, sep = '-') %>%
  dplyr::select(-Comparison) %>%
  dplyr::rename(Region = BrodmannArea, Comparison = Dx, gene.id = genes) %>%
  dplyr::mutate(Region = factor(Region, labels = c('FP', 'STG', 'PHG', 'IFG')),
                Study = 'MSBB') %>%
  left_join(backgroundGenes) %>%
  left_join(Ensemble2HGNC) %>%
  dplyr::select(-position) %>%
  dplyr::mutate(Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction))
DE1$Direction[DE1$adj.P.Val > 0.05 | abs(DE1$logFC) < log2(1.2)] = 'NONE'

p = ggplot(DE1, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red'))
p = p + facet_grid(Region~.+Comparison, scales = 'fixed')
p

@@

### Associate differential expression results with gc content and gene length
```{r associate.de}
geneLength.GCcontent = EDASeq::getGeneLengthAndGCContent(unique(DE1$ensembl_gene_id), 
                                                         org = 'hsa', 
                                                         mode = "biomart") %>%
  rownameToFirstColumn('ensembl_gene_id')

tmp = left_join(DE1, geneLength.GCcontent)

pl = list()
pl[[1]] = ggplot(tmp, aes(y = log10(length), x = abs(logFC))) + 
  geom_point() + 
  stat_smooth(method=glm, se = TRUE, na.rm = T)
pl[[2]] = ggplot(tmp, aes(y = gc, x = abs(logFC))) + 
  geom_point() + 
  stat_smooth(method=glm, se = TRUE, na.rm = T)
pl[[3]] = ggplot(tmp, aes(y = log10(length), x = adj.P.Val)) + 
  geom_point() + 
  stat_smooth(method=glm, se = TRUE, na.rm = T)
pl[[4]] = ggplot(tmp, aes(y = gc, x = adj.P.Val)) + 
  geom_point() + 
  stat_smooth(method=glm, se = TRUE, na.rm = T)

multiplot(plotlist = pl, cols = 2)
```

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Add gene names
RESIDUAL.GENE_EXPRESSION <- RESIDUAL.GENE_EXPRESSION %>%
  rownameToFirstColumn('gene.id') %>%
  left_join(backgroundGenes) %>%
  left_join(Ensemble2HGNC) %>%
  dplyr::select(gene.id, ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))

VOOM.ADJUSTED.GENE_EXPRESSION$E <- VOOM.ADJUSTED.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('gene.id') %>%
  left_join(backgroundGenes) %>%
  left_join(Ensemble2HGNC) %>%
  dplyr::select(gene.id, ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))

# Code
CODE <- Folder(name = 'MSBB Reprocessed RNASeq TMM Binned CDR', parentId = parentId)
annotations(CODE) = list(
  dataType = 'mRNA',
  dataSubType = 'geneExp',
  summaryLevel = 'gene',
  assay	 = 'Reprocessed RNAseq',

@@

# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'sampleIdentifier')
write.table(COVARIATES, file = 'MSBB_Covariates.tsv', sep = '\t', row.names=F, quote=F)
COV_OBJ = File('MSBB_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
annotations(COV_OBJ) = annotations(CODE)
annotations(COV_OBJ)$dataSubType = 'covariates'
COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix %>%
  as.matrix() %>% rownameToFirstColumn('gene.id') %>%
  write.table(file = 'MSBB_Counts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('MSBB_Counts.tsv', name = 'Raw counts (filtered)', parentId = CODE$properties$id)
annotations(COUNT_OBJ) = annotations(CODE)
annotations(COUNT_OBJ)$dataSubType = 'filteredCounts'
COUNT_OBJ = synStore(COUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                     executed = thisFile, activityDescription = activityDescription)

# Store logCPM
write.table(VOOM.ADJUSTED.GENE_EXPRESSION$E, file = 'MSBB_logCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('MSBB_logCPM.tsv', name = 'Raw counts (logCPM)', parentId = CODE$properties$id)
annotations(LCOUNT_OBJ) = annotations(CODE)
annotations(LCOUNT_OBJ)$dataSubType = 'logcpm'
LCOUNT_OBJ = synStore(LCOUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)

# Store design matrix
write.table(DM1$design, file = 'MSBB_Design.tsv', sep = '\t', row.names=F, quote=F)
DM_OBJ = File('MSBB_Design.tsv', name = 'Design Matrix', parentId = CODE$properties$id)
annotations(DM_OBJ) = annotations(CODE)
annotations(DM_OBJ)$dataSubType = 'designMatrix'
DM_OBJ = synStore(DM_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                  executed = thisFile, activityDescription = activityDescription)

# Store residual gene expression
write.table(RESIDUAL.GENE_EXPRESSION, file = 'MSBB_Expression.tsv', sep = '\t', row.names=F, quote=F)
EXP_OBJ = File('MSBB_Expression.tsv', name = 'Normalised Adjusted Residual Expression (Dx added)', 
               parentId = CODE$properties$id)
annotations(EXP_OBJ) = annotations(CODE)
annotations(EXP_OBJ)$dataSubType = 'residualGeneExp'
EXP_OBJ = synStore(EXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)

# Store differential expression results
write.table(DE1, file = 'MSBB_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('MSBB_DiffExpression.tsv', name = 'Differential Expression Analysis', parentId = CODE$properties$id)
annotations(DEXP_OBJ) = annotations(CODE)
annotations(DEXP_OBJ)$dataSubType = 'diffExp'
DEXP_OBJ = synStore(DEXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                    executed = thisFile, activityDescription = activityDescription)
stopCluster(cl)
```
|  *Results*                |  *SynapseID*                    |
|  -------                  |   ---------                     |
|  Covariates               |  `r COV_OBJ$properties$id`      |
|  Raw counts (logCPM)      |  `r LCOUNT_OBJ$properties$id`   |
|  Design Matrix            |  `r DM_OBJ$properties$id`       |
|  Residual Expression      |  `r EXP_OBJ$properties$id`      |
|  Differential Expression  |  `r DEXP_OBJ$properties$id`     |

### 2017-03-07 code/Rmd/ROSMAP_RNASeq_Reprocessed_TMM.Rmd
Added picard metrics to MSSM reprocessed rnaseq analysis

### Covariate clustering
Determine relationship between covariates
```{r covariates.clustering, cache=TRUE}
#"braaksc","ceradsc","cts_mmse30_first_ad_dx","cts_mmse30_lv", "PF_INDEL_RATE", "PF_MISMATCH_RATE"
FactorCovariates <- c("Batch", "msex", "race", "spanish", "cogdx")
ContCovariates <- c("RINcontinuous","age_first_ad_dx","age_death","age_at_visit_max","pmi","educ", 
                    "PCT_PF_READS_ALIGNED", "PCT_CORRECT_STRAND_READS", "PCT_INTERGENIC_BASES", 
                    "PCT_INTRONIC_BASES", "PCT_RIBOSOMAL_BASES", "PCT_USABLE_BASES")

>>>>

### Covariate clustering
Determine relationship between covariates
```{r covariates.clustering, cache=TRUE}
#"braaksc","ceradsc","cts_mmse30_first_ad_dx","cts_mmse30_lv", "PF_INDEL_RATE", "PF_MISMATCH_RATE"
FactorCovariates <- c("Batch", "msex", "race", "spanish", "cogdx")
ContCovariates <- c("RINcontinuous","age_first_ad_dx","age_death","age_at_visit_max","pmi","educ", 
                    "PCT_PF_READS", "PCT_PF_READS_ALIGNED", "PCT_INTERGENIC_BASES", "PCT_INTRONIC_BASES", 
                    "PCT_RIBOSOMAL_BASES", "PCT_USABLE_BASES")
